---
title: "HLAexpression"
author: "Cyril Pottier"
date: "2/8/2021"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/tijs/Documents/RademakersLab/R')

```



```{r}
setwd("C:/Users/tijs/Documents/RademakersLab/R")


#Loading libraries
library(stringi)
library(cqn)
library(emmeans)
library(data.table)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(gtools)
library('biomaRt')
library("EnhancedVolcano")
library("anRichment")
library(tibble)
library(dplyr)

#BiocManager::install("EnhancedVolcano")
```

# Prepare the dataset
```{r}
options( stringsAsFactors = FALSE)

#(file = paste("run_",format(Sys.time(),"%Y%m%d-%Hu%M"),"_lm.log",sep=""), append = TRUE, split=TRUE,type = "output")
counts<-data.frame(read.csv("RNAseqCounts.csv",header=T,sep=";"))
sampleinfo<-data.frame(read.csv("RNAseqSample_info.txt",header=T,sep="\t"))
sampleinfo<-subset(sampleinfo,Age!="NA") #this removes one sample without AAD

samplealiases<-read.csv("sample_aliases.txt",header=F,sep="\t")
#Rename the columns in samplealisaes file
names(samplealiases)<-c("seqid","sampleid")

#Change the sample IDs
samplealiases$sampleid<-sapply(stri_split_fixed(samplealiases$sampleid, '-', n=2),'[',2)

#replace - with . in the seqi column from the samplealiases
samplealiases$seqid<-gsub("-", ".", samplealiases$seqid)

#remove the _count characters in the counts df
names(counts)<-gsub("_count", "", names(counts))

#Create a vector of character with the sampleID
seqnames<- names(counts)

#Put the plate number in the plate column of the sampleinfo df
sampleinfo$plate<-sapply(stri_split_fixed(sampleinfo$plate, '-', n=3),'[',2)

counts<-data.frame(read.csv("RNAseqCounts.csv",header=T,sep=";"))
names(counts) <- sub("X.home.MOLGEN.twatzeels.app.featureCounts.results.BAM_20210207.", "", names(counts))
names(counts) <- sub(".bam", "", names(counts))


#Replace the sample ID in the counts df with the actual STX ID using the samplealiases file
for(i in 7:ncol(counts)){
  for(j in 1:nrow(samplealiases)){
      if(names(counts)[i]==samplealiases[j,1]){
      names(counts)[i]=samplealiases[j,2];
	}    
    }
}
  

# # rename the groups according to the correct column
 sampleinfo$groups[sampleinfo$GROUP=='FTLD-TDP-A']<-"TDP"
 sampleinfo$groups[sampleinfo$GROUP=='FTLD-TDP-B']<-"TDP"
 sampleinfo$groups[sampleinfo$GROUP=='FTLD-TDP-C']<-"TDP"
 sampleinfo$groups[sampleinfo$GROUP=='Control']<-"Control"

#Selecton only the TDP and Controls
sampleinfo<-sampleinfo[ grep("TDP|Control",sampleinfo$groups), ]

#Subset the sampleinfo file
sampleinfo<-sampleinfo[,c("SequencingID","Age","Sex", "groups","RINFCX","plate","GROUP")] 

#Create a list of samples actually used (with data)
used_samples<-intersect(sampleinfo$SequencingID,names(counts))

#Subset the counts df into the samples that we actually want to use
countsraw<- counts[,used_samples] 
nrow(countsraw)

toadd<-read.table("merged_gene_count.csv",header=T,sep=";")

toadd<- toadd[,1:7]
library(data.table)
test<-merge(counts, toadd, by.x="Geneid", by.y="GeneId")

#Get the annotation of the transcripts from the counts file
annotcountsraw<-test[,c(1:6, ncol(test))]

#Add the geneId column into the rownames of the countsraw df
rownames(countsraw)<-counts$Geneid
######Now can get rid of the counts df and pursue with the countsraw###################################################


#Clean up unecessary files
#rm(counts, samplealiases, j, i, seqnames, used_samples)
```

## CQN normalization
```{r}
### cqn normalization 
#geneinfo<-data.frame(read.csv("gene_info_gc_length_clean.tsv",header=T,sep="\t"))

geneinfo<- read.table("Ensembl_hg38.Gene_features.GC_Content.txt",sep="\t", header=TRUE, stringsAsFactors=FALSE)

# Remove Chr, Start and End column
annotcountsraw <- annotcountsraw[,-c(2:5)]

# Merge the annotation matrix to get the length and the GC conent for all genes
geneinfocomplete<-merge(annotcountsraw, geneinfo, by.x="Geneid", by.y="GeneID")

#Subest the geneinfocomplete df to the transcripts we have
geneinfocomplete<-geneinfocomplete[match(rownames(countsraw), geneinfocomplete$Geneid),]

#Clean up some df and lists
#rm(annotcountsraw)
```

## Identify individuals with issues
```{r}
indv_problem<-c()
 for(i in 1:ncol(countsraw)){
 Male<-max(countsraw[which(row.names(countsraw)=="ENSG00000129824"),i],countsraw[which(row.names(countsraw)=="ENSG00000198692"),i],countsraw[which(row.names(countsraw)=="ENSG00000067048"),i],countsraw[which(row.names(countsraw)=="ENSG00000012817"),i])
 Female<-countsraw[which(row.names(countsraw)=="ENSG00000229807"),i]
 ratio=(Male+1)*100/(Male+Female+1)
#  print(paste("Wrong gender assigned!!", names(annotcountsraw)[i], Male,Female,ratio, sep="\t"))
    if(ratio>75.0 & sampleinfo[which(sampleinfo$SequencingID ==names(countsraw)[i]),3]=="F"){print(paste("Wrong gender assigned!! Indv is not a female", names(countsraw)[i], Male,Female,ratio, sep="\t"));
    indv_problem<- c(indv_problem, names(countsraw)[i])
    }
    if(ratio<20.0 & sampleinfo[which(sampleinfo$SequencingID ==names(countsraw)[i]),3]=="M"){print(paste("Wrong gender assigned!!Indv is not a male", names(countsraw)[i], Male,Female,ratio, sep="\t"));
    indv_problem<- c(indv_problem, names(countsraw)[i])
 }
    }

indv_problem<-c(indv_problem,as.character(sampleinfo[which(is.na(sampleinfo$RINFCX)),c("SequencingID")]))
indv_problem<-c(indv_problem,as.character(sampleinfo[which(is.na(sampleinfo$Age)),c("SequencingID")]))
indv_problem
```

### Remove the individuals with sex discrepencies
```{r}
countsraw<-countsraw[ , -which(names(countsraw) %in% indv_problem)]
sampleinfo<-sampleinfo[ ! sampleinfo$SequencingID %in% indv_problem, ]

#Clean-up 
#rm(geneinfo, Female, Male, i, indv_problem, ratio)
```


### Remove individuals with GC content of zero
#```{r}
#geneinfocomplete<-geneinfocomplete[!(geneinfocomplete$GC_Content==0),]
#countsraw<-countsraw[match(geneinfocomplete$Geneid,rownames(countsraw)),]
#```

### Remove any NA observations in geneinfo dataset?
```{r}
#omit NA values
geneinfocomplete<-na.omit(geneinfocomplete)

#re-subset the dataframes
countsraw<-subset(countsraw, rownames(countsraw) %in% geneinfocomplete$Geneid)
```


### Actual CQN
```{r}
set.seed(101)
cqn<- cqn(countsraw, lengths = geneinfocomplete$Length, x = geneinfocomplete$GC_Content,  verbose = TRUE)
#See the man page for ?cqn under the description of the output glm.offset (cqn corrects for size factors and the glm.offset contains size factor correction)
#cnq.sizefactors<-cqn$glm.offset

###We can examine plots of systematic effects by using cqnplot. The n argument refers to the systematic effect, n=1 is always the covariate specified by the x argument above, while n=2 is lengths.
par(mfrow=c(1,2))
cqnplot(cqn, n = 1, xlab = "GC content", lty = 1, ylim = c(1,7))
cqnplot(cqn, n = 2, xlab = "length", lty = 1, ylim = c(1,7))

#The normalized expression values are (These values are on the log2-scale)
RPKM.cqn_all <- cqn$y + cqn$offset
RPKM.cqn_all[1:4,1:4]


#we need to use the component glm.offset which is on the natural logarithmic scale and also includes correcting for sizeFactors. It is possible to include the offset directly into the DGEList, by post-processing the output like:

cqnGlm <- cqn$glm.offset #This is to be used as input for EdgeR or DEseq2. cqn$y + cqn$offset is a CQN-corrected estimated expression measures (not on the log scale)
cqnGlm[1:4,1:4]

save(RPKM.cqn_all, cqnGlm, file="Gene_Count_CQN.RData")
write.table(RPKM.cqn_all,"RPKM_cqn.tsv",quote = FALSE, sep = "\t",na = "NA", dec = ".", row.names = TRUE,col.names=NA)

#filter out the normalized counts based on the maximum value in each row of my data, to be greater than 0
RPKM.cqn_all<-data.frame(RPKM.cqn_all,check.names=FALSE)
rownames(RPKM.cqn_all)<-rownames(countsraw)


RPKM.cqn<-RPKM.cqn_all

annot_RPKM.cqn<-cbind(geneinfocomplete,RPKM.cqn_all)

RPKM.cqn<-annot_RPKM.cqn[apply(annot_RPKM.cqn[,9:ncol(annot_RPKM.cqn)], 1, function(x) (max(x) > 0)), ]
nrow(RPKM.cqn)


# Remove the non-protein coding genes and the X or Y chromosome genes
RPKM.cqn_prot<-RPKM.cqn[which(RPKM.cqn$GeneBiotype=="protein_coding"),]
nrow(RPKM.cqn_prot)
RPKM.cqn_prot<-RPKM.cqn_prot[-grep("chrY",RPKM.cqn_prot$Chr),]
nrow(RPKM.cqn_prot)
RPKM.cqn_prot<-RPKM.cqn_prot[-grep("chrM",RPKM.cqn_prot$Chr),]
nrow(RPKM.cqn_prot)
RPKM.cqn_prot<-RPKM.cqn_prot[-grep("CHR_",RPKM.cqn_prot$Chr),]
nrow(RPKM.cqn_prot)

row.names(RPKM.cqn_prot)<-RPKM.cqn_prot$GeneId
##ENO2 ENSG00000111674
##CD68 ENSG00000129226
##CD34 ENSG00000174059
##GFAP ENSG00000131095
##OLIG2 ENSG00000205927

#Clean-up
#rm(RPKM.cqn, RPKM.cqn_all,cqn,annot_RPKM.cqn, countsraw,cqnGlm, geneinfocomplete)
```



# Differential Expression code and generation of summary file
```{r}
groups <- factor(sampleinfo$groups, levels = c("Control","TDP"))
sex<-factor(sampleinfo$Sex, levels = c("M","F"))
plate<-factor(sampleinfo$plate, levels = c("Plate1", "Plate2"))



# Function to calculate fold change for variable A (a1) and B (b1)

foldChange <- function(a1, b1){
  if(a1 > b1){
    2^(a1 - b1)
  }
  else{
    -(2^(b1 - a1))
  }
}


###no model adjustment
pValues <- c()
rValues <- c()

for(i in 1:nrow(RPKM.cqn_prot)){
expr<-t(RPKM.cqn_prot[i,9:ncol(RPKM.cqn_prot)])
f1<-lm(expr~sampleinfo$RINFCX+sampleinfo$Sex+sampleinfo$Age+sampleinfo$plate+groups, na.action=na.exclude)

f1nogroup<-lm(expr~sampleinfo$RINFCX+sampleinfo$Sex+sampleinfo$Age+sampleinfo$plate, na.action=na.exclude)

a1<-anova(f1)  
## Extract p-values for variables
  pValues$GeneId[i] <- RPKM.cqn_prot[i,"Geneid"]
  pValues$symbol[i]<-RPKM.cqn_prot[i,"GeneName"]
  pValues$PValueGroup[i] <- a1$"Pr(>F)"[5]
  pValues$PValueRINFCX[i] <- a1$"Pr(>F)"[1]
  pValues$PValueSex[i] <- a1$"Pr(>F)"[2]
  pValues$PValueAge[i] <- a1$"Pr(>F)"[3]
  pValues$PValuePlate[i] <- a1$"Pr(>F)"[4]
  
## Extract emmeans
### Make reference grid first

  ref1 <- ref_grid(f1) 
  
  ls1 <- summary(emmeans(ref1, "groups"))
  pValues$EMMeanControl[i] <- ls1$emmean[1]
  pValues$EMMeanTDP[i] <- ls1$emmean[2]
  
## Extract p-values for group variable and add fold change
### A tukey method is used for comparing a family of estimates

  ls2 <- summary(emmeans(ref1, pairwise ~ groups))
  
   pValues$PValueTDP_vs_Control[i] <- ls2$contrasts$p.value[1]
  pValues$FoldChangeTDP_vs_Control[i] <- foldChange(ls1$emmean[2], ls1$emmean[1])
  
  
  ## Extract residuals

  r1 <- residuals(f1nogroup)
  rValues$i <- r1
  names(rValues)[i] <- RPKM.cqn_prot[i, "Geneid"]
  
#names(rValues)[i] <- row.names(RPKM.cqn[i,])

#  residmat<-data.frame(resid(f1nogroup),groups)
# names(residmat)<-c("expr_residuals","group")
# # pdf("_wo_correction.pdf")
# png("_wo_correction.png")
#   p <- ggplot(residmat,aes(x=group, y=expr_residuals,fill=group)) + ggtitle("\n(without cell type markers)")+geom_boxplot()+ geom_jitter(shape=16, position=position_jitter(0.2))
#  print(p)
#  dev.off()
#  
}



#Create the file with the pvalues and fold changes
pValues$PValueTDP_vs_Control_fdr<-p.adjust(pValues$PValueTDP_vs_Control, method = "fdr", n = length(pValues$PValueTDP_vs_Control))

pValues$PValueTDP_vs_Control_bon<-p.adjust(pValues$PValueTDP_vs_Control, method = "bonferroni", n = length(pValues$PValueTDP_vs_Control))

f1_output<-data.frame(pValues)

f1_output<-f1_output[,c("GeneId","symbol","PValueGroup", "PValueRINFCX", "PValueSex", "PValueAge","PValuePlate","EMMeanTDP","EMMeanControl","FoldChangeTDP_vs_Control","PValueTDP_vs_Control","PValueTDP_vs_Control_fdr","PValueTDP_vs_Control_bon")]

write.table(f1_output,"FCX_Table_FC_pval.tsv",quote = FALSE, sep = "\t",na = "NA", dec = ".", row.names = TRUE,col.names=NA)


residf1_output<-data.frame(rValues)
residf1_table<-transpose(residf1_output)
row.names(residf1_table)<-names(residf1_output)
names(residf1_table)<-row.names(residf1_output)

residf1_table$symbol = mapIds(org.Hs.eg.db,keys=as.vector(row.names(residf1_table)), column="SYMBOL",keytype="ENSEMBL",multiVals="first")
write.table(residf1_table,"FCX_Table_residuals.tsv",quote = FALSE, sep = "\t",na = "NA", dec = ".", row.names = TRUE,col.names=NA)


```

```{r}
#Clean-up
#rm(list= ls()[!(ls() %in% c('f1_output','RPKM.cqn_prot', 'sampleinfo'))])
```


## FTLD-TDP vs controls
### Prepare the data for plotting
```{r}
grouping<-"TDPvsControls"
directory<-"./grouping"
dir.create(directory)
Expression_data_all<- f1_output
#Expression_data_all <- na.omit(Expression_data_all)

grep("^FoldChangeTDP_vs_Control$", colnames(Expression_data_all))

colnames(Expression_data_all)

Expression_data<-Expression_data_all[, c(1,2,10:13)]
Expression_data$Log2FC<-foldchange2logratio(Expression_data$FoldChangeTDP_vs_Control, base=2)
names(Expression_data)[names(Expression_data) == "PValueTDP_vs_Control_fdr"] <- "Pvalue_fdr"
names(Expression_data)[names(Expression_data) == "PValueTDP_vs_Control_bon"] <- "Pvalue_bon"
names(Expression_data)[names(Expression_data) == "PValueTDP_vs_Control"] <- "Pvalue"
names(Expression_data)[names(Expression_data) == "GeneId"] <- "GeneSymbol"
names(Expression_data)[names(Expression_data) == "FoldChangeTDP_vs_Control"] <- "FoldChange"
mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
genes <- Expression_data$GeneSymbol
G_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id","hgnc_symbol"),values=genes,mart= mart)
Expression_data<-merge(Expression_data,G_list,by.x="GeneSymbol",by.y="ensembl_gene_id")

#Clean-up
#rm(G_list, mart,genes)
```

###Extract the names of the HLA genes present in the Expression dataset
```{r}

Gene_names <- Expression_data[, c("symbol"), drop=FALSE]
HLA_list <- Gene_names[ c(grep("HLA-", Gene_names$symbol)), ]
HLA_names <- data.frame(matrix(unlist(HLA_list), nrow=length(HLA_list), byrow=TRUE))
head(HLA_names)


```



```{r}
geom_label_repel(
  max.overlaps = 200000
)


 EnhancedVolcano(Expression_data,
    lab = as.character(Expression_data$symbol),
    x = 'Log2FC',
    y = 'Pvalue',
    xlim = c(min(Expression_data[["Log2FC"]], na.rm = TRUE) - 0.5, max(Expression_data[["Log2FC"]], na.rm = TRUE) +
0.5),
    ylim = c(0, max(-log10(Expression_data[["Pvalue"]]), na.rm = TRUE) + 1),
    axisLabSize = 10.0,
    title = 'Differential Gene expression analysis',
    titleLabSize = 13.0,
    subtitle = 'FTLD-TDP vs Controls',
    subtitleLabSize = 10.0,
    caption = bquote('A Volcanoplot showing the differences in gene expression 
                     when comparing FTLD-TDP patients to a control group.'),
    captionLabSize = 9.0,
    pCutoff = 10e-8,
    FCcutoff = 1.5,
    pointSize = 1.0,
    selectLab = HLA_names$matrix.unlist.HLA_list...nrow...length.HLA_list...byrow...TRUE.,
    labSize = 3.0,
    labCol = "red2",
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    col=c("grey20", "royalblue", "forestgreen", "red2" ),
    colAlpha = 0.65,
    legendLabels=c('No significance',expression(Log[2]~FC), expression(italic (p-value)),  expression(italic(p-value)~and~Log[2]~FC)),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 3.0,
    gridlines.major = TRUE,
    gridlines.minor = TRUE,
    border = "full",
    borderWidth = 0.5)
```


```{r}

 EnhancedVolcano(Expression_data,
    lab = as.character(Expression_data$symbol),
    x = 'Log2FC',
    y = 'Pvalue',
    axisLabSize = 10.0,
    title = 'Differential Gene expression analysis',
    titleLabSize = 13.0,
    subtitle = 'FTLD-TDP vs Controls',
    subtitleLabSize = 10.0,
    caption = bquote('A Volcanoplot showing the differences in gene expression 
                     when comparing FTLD-TDP patients to a control group.'),
    captionLabSize = 9.0,
    pCutoff = 10e-8,
    FCcutoff = 1.5,
    pointSize = 1.0,
    labSize = 3.0,
    labCol = "red2",
    col=c("grey20", "royalblue", "forestgreen", "red2" ),
    colAlpha = 0.65,
    legendLabels=c('No significance',expression(Log[2]~FC), expression(italic (p-value)),  expression(italic(p-value)~and~Log[2]~FC)),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 3.0,
    gridlines.major = TRUE,
    gridlines.minor = TRUE,
    border = "full",
    borderWidth = 0.5)

```


```{r}
 EnhancedVolcano(Expression_data,
    lab = as.character(Expression_data$symbol),
    x = 'Log2FC',
    y = 'Pvalue_fdr',
    xlim = c(min(Expression_data[["Log2FC"]], na.rm = TRUE) - 3.0, max(Expression_data[["Log2FC"]], na.rm = TRUE) +
3.0),
    ylim = c(0, max(-log10(Expression_data[["Pvalue_fdr"]]), na.rm = TRUE) + 1),
    axisLabSize = 10.0,
    title = 'Differential Gene expression analysis',
    titleLabSize = 13.0,
    subtitle = 'FTLD-TDP vs Controls',
    subtitleLabSize = 10.0,
    caption = bquote('Adjusted p-values (fdr) were used'),
    captionLabSize = 9.0,
    pCutoff = 0.05,
    FCcutoff = 1.0,
    pointSize = 0.1,
    labSize = 2.0,
    labCol = "black",
    #drawConnectors = TRUE,
    #widthConnectors = 0.5,
    col=c("grey20", "royalblue", "forestgreen", "red2" ),
    colAlpha = 0.75,
    legendLabels=c('No significance',expression(Log[2]~FC), expression(italic (p-value)),  expression(italic(p-value)~and~Log[2]~FC)),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 3.0,
    gridlines.major = TRUE,
    gridlines.minor = TRUE,
    border = "full",
    borderWidth = 0.5)
```

```{r}
 EnhancedVolcano(Expression_data,
    lab = as.character(Expression_data$symbol),
    x = 'Log2FC',
    y = 'Pvalue_bon',
    xlim = c(min(Expression_data[["Log2FC"]], na.rm = TRUE) - 3.0, max(Expression_data[["Log2FC"]], na.rm = TRUE) +
3.0),
    ylim = c(0, max(-log10(Expression_data[["Pvalue_bon"]]), na.rm = TRUE) + 1),
    axisLabSize = 10.0,
    title = 'Differential Gene expression analysis',
    titleLabSize = 13.0,
    subtitle = 'FTLD-TDP vs Controls',
    subtitleLabSize = 10.0,
    caption = bquote('Adjusted p-values (bonferroni) were used'),
    captionLabSize = 9.0,
    pCutoff = 0.05,
    FCcutoff = 0.58496250072,
    pointSize = 0.1,
    labSize = 2.0,
    labCol = "black",
    #drawConnectors = TRUE,
    #widthConnectors = 0.5,
    col=c("grey20", "royalblue", "forestgreen", "red2" ),
    colAlpha = 0.75,
    legendLabels=c('No significance',expression(Log[2]~FC), expression(italic (p-value)),  expression(italic(p-value)~and~Log[2]~FC)),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 3.0,
    gridlines.major = TRUE,
    gridlines.minor = TRUE,
    border = "full",
    borderWidth = 0.5)
```


```{r}
 EnhancedVolcano(Expression_data,
    lab = as.character(Expression_data$symbol),
    x = 'Log2FC',
    y = 'Pvalue_fdr',
    xlim = c(min(Expression_data[["Log2FC"]], na.rm = TRUE) - 0.5, max(Expression_data[["Log2FC"]], na.rm = TRUE) +
0.5),
    ylim = c(0, max(-log10(Expression_data[["Pvalue_fdr"]]), na.rm = TRUE) + 1),
    axisLabSize = 10.0,
    title = 'Differential Gene expression analysis',
    titleLabSize = 13.0,
    subtitle = 'FTLD-TDP vs Controls',
    subtitleLabSize = 10.0,
    caption = bquote('Adjusted p-values were used'),
    captionLabSize = 9.0,
    pCutoff = 0.05,
    FCcutoff = 0.58496250072, #this is the value for log2(1.5), because 1.5 is a cutoff used at linear FC
    pointSize = 0.4,
    labSize = 2.0,
    labCol = "red2",
    #drawConnectors = TRUE,
    #widthConnectors = 0.5,
    colGradient = c('red3', 'royalblue'),
    colAlpha = 0.75,
    legendLabels=c('No significance',expression(Log[2]~FC), expression(italic (p-value)),  expression(italic(p-value)~and~Log[2]~FC)),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 3.0,
    gridlines.major = TRUE,
    gridlines.minor = TRUE,
    border = "full",
    borderWidth = 0.5)
```

