---
title: "LolliPopPlot"
output: html_notebook
---

```{r}

###SetupDir
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/tijs/Documents/RademakersLab/R')

```

###load libraries
```{r}
setwd("C:/Users/tijs/Documents/RademakersLab/R")

#Loading libraries
library(stringi)
library(data.table)
library('biomaRt')
library(RColorBrewer)
library(tidyverse)

library(Gviz)
library(rtracklayer)
library(trackViewer)

```

###LoliPopPlot
```{r}
#SNP locations
SNP<-c(3332, 3380, 3991)
#IDUNNOOYET
gr<-GRanges("chr6", IRanges(SNP, width = 1, names = paste0("SNP", SNP)))
#Coding regions
feat<- GRanges("chr6", IRanges(c(65, 2526, 3299, 3870),
                               width = c(81, 281, 281, 154),
                               names = paste0("exon", 1:4)))
lolliplot(gr, feat)

feat$fill <- "#51C6E6"
gr$color <- c("#1F78B4", "#1F78B4", "#E31A1C")
gr$border <- sample(("gray30"), length(SNP), replace=TRUE)
leg<-c("silent", "silent", "missense")
gr$leg<- leg

lolliplot(gr, feat, legend = "leg")


gr.rot <- gr
gr.rot$label.parameter.rot <- 45
lolliplot(gr.rot, feat, legend="leg")

#rs7192, also known as Leu242Val, G>T, position 32443869 (3991).
display.brewer.all(colorblindFriendly = T)
brewer.pal(n=9, name = "Paired")
display.brewer.pal(n=9, name = "Paired")

```





