---
title: "PlotHLAgenesSeparated"
output: html_notebook
---

```{r, setup}
setwd("C:/Users/tijs/Documents/Rademakers Lab")
```

###set work directory
#setwd("C:/Users/tijs/Documents/Rademakers Lab")

#knitr::opts_knit$set(root.dir = #'C:/Users/tijs/Documents/Rademakers Lab')


```{r}
library(tidyverse)
library(ggplot2)
library(stringr)
library(data.table)
```


###If you want to make a separate control and tdp dataset for plotting
#we won't use it now but I spent a lot of time making it, so I cannot bring myself to throw it away yet
```{r}
#Make two separate dataframes, one for controls and one for patients. 
#Transpose them and then change - by . to match platenames
#sampleinfocontrol<-sampleinfo[sampleinfo$GROUP == "Control",]
#sampleinfotdp<-sampleinfo[sampleinfo$groups == "TDP",]
#sampleinfocontrol<-sampleinfocontrol[,c(1,4)]
#sampleinfotdp<-sampleinfotdp[,c(1,4)]
#sampleinfocontrol$SequencingID<-gsub("-",".",sampleinfocontrol$SequencingID)
#sampleinfotdp$SequencingID<-gsub("-",".",sampleinfotdp$SequencingID)

#Not sure if this is a necessary step, but I believe it is to later be able to use the "select" option (tidyverse)
#test<- as_tibble(SubsetC)


#☻This gives you the list for the select-option in the next step
#cat(paste(shQuote(sampleinfocontrol, type="cmd"), collapse=", "))
#cat(paste(shQuote(sampleinfotdp, type="cmd"), collapse=", "))


#residcon<-test %>% select("hgnc_symbol", "NA99.013", "NA05.343", "NA98.318", "NA03.010", "NA02.098", "NA98.350", "NA03.248", "NA09.354", "NA04.127", "NA05.131", "X5138", "JA96.053", "NA04.006", "NA07.279", "NA98.268", "NA98.267", "NA02.048", "NA07.215", "NA98.274", "NA07.238", "NA09.421", "NA05.055", "NA98.356", "NA12.140")
#Attention! somewhere platenames that exist out of only numbers (e.g. 5138) got changed
#To e.g. X5138... I don't know what caused this

#residtdp<-test %>% select("hgnc_symbol","NA13.069", "NA00.217", "NA11.427", "NA09.119", "NA11.033", "NA06.094", "NA09.204", "NA04.327", "NA98.230", "NA06.149", "NA05.122", "NA02.001", "NA10.324", "NA09.024", "NA11.013", "NA05.237", "X4706", "NA08.273", "NA06.199", "NA10.349", "NA02.194", "NA09.018", "X3970", "NA04.145", "NA10.336", "NA09.369", "NA11.305", "NA10.188", "NA12.150", "NA12.258", "NA02.045", "NA11.395", "NA12.350", "NA10.147", "NA00.138", "NA98.175", "NA11.015", "NA01.013", "NA05.367", "NA09.392", "NA11.212", "NA10.407", "NA09.312", "NA11.486", "NA05.065", "NA02.049", "NA05.324", "NA00.025", "NA04.111", "NA11.195", "NA10.421", "NA02.276", "NA08.051", "NA10.049", "NA98.195", "NA04.167", "NA00.186", "NA01.123", "NA98.204", "NA00.086", "NA11.201", "NA10.303", "NA02.270", "NA09.031", "NA02.144")
#Here, 4706 and 3970 were also changed manually to X4706 and X3970 to match the platename in SubsetC


#use the column containing gene names as rownames instead
#residcon<-column_to_rownames(residcon, 'hgnc_symbol')
#residtdp<-column_to_rownames(residtdp, 'hgnc_symbol')


#Now you have a subset containing the residuals for the controls and the patients, respectively
#Cleanup
#rm (SubsetA, SubsetB, SubsetC, test, num)
```


###Preparation of the data
```{r}

test<-residf1_output

#subdoodle<-test[, "ENSG00000206503"]
#rm(subdoodle)

#grab only the HLA genes from the expression output
SubsetA<-test[,c("ENSG00000206503","ENSG00000234745", "ENSG00000204525", "ENSG00000196735", "ENSG00000179344", "ENSG00000237541", "ENSG00000232629", "ENSG00000204287", "ENSG00000196126", "ENSG00000198502", "ENSG00000231389", "ENSG00000223865", "ENSG00000204592", "ENSG00000204642", "ENSG00000204632", "ENSG00000204257", "ENSG00000204252", "ENSG00000241106", "ENSG00000242574")]

#What about DPA2 and DPB2?
#,"ENSG00000231461","ENSG00000224557"

SubsetA<-data.frame(SubsetA, subsampleinfo)


```


###HLA-A
```{r}
#Make plot for HLA-A expression, Comparing patient and control samples
#ENSG00000206503

#Grab HLA-A from the subset
scplot<-SubsetA[,c("ENSG00000206503","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000206503,)) + 
  ylab("HLA-A")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )

#different option for plotting, uses ggplot (better)


```

###HLA-B
```{r}
#ENSG00000234745
scplot<-SubsetA[,c("ENSG00000234745","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000234745,)) + 
  ylab("HLA-B")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-C
```{r}
#ENSG00000204525
scplot<-SubsetA[,c("ENSG00000204525","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000204525,)) + 
  ylab("HLA-C")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DQA1
```{r}
#ENSG00000196735
scplot<-SubsetA[,c("ENSG00000196735","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000196735,)) + 
  ylab("HLA-DQA1")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DQB1
```{r}
#ENSG00000179344
scplot<-SubsetA[,c("ENSG00000179344","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000179344,)) + 
  ylab("HLA-DQB1")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DQA2
```{r}
#ENSG00000237541
scplot<-SubsetA[,c("ENSG00000237541","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000237541,)) + 
  ylab("HLA-DQA2")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )



```

###HLA-DQB2
```{r}
#ENSG00000232629
scplot<-SubsetA[,c("ENSG00000232629","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000232629,)) + 
  ylab("HLA-DQB2")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )



```

###HLA-DRA
```{r}
#ENSG00000204287
scplot<-SubsetA[,c("ENSG00000204287","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000204287,)) + 
  ylab("HLA-DRA")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DRB1
```{r}
#ENSG00000196126
scplot<-SubsetA[,c("ENSG00000196126","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000196126,)) + 
  ylab("HLA-DRB1")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DRB5
```{r}
#ENSG00000198502
scplot<-SubsetA[,c("ENSG00000198502","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000198502,)) + 
  ylab("HLA-DRB5")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DPA1
```{r}
#ENSG00000231389
scplot<-SubsetA[,c("ENSG00000231389","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000231389,)) + 
  ylab("HLA-DPA1")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DPB1
```{r}
#ENSG00000223865
scplot<-SubsetA[,c("ENSG00000223865","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000223865,)) + 
  ylab("HLA-DPB1")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

#What about DPA2 and DPB2?


###HLA-E
```{r}
#ENSG00000204592
scplot<-SubsetA[,c("ENSG00000204592","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000204592,)) + 
  ylab("HLA-E")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-F
```{r}
#ENSG00000204642
scplot<-SubsetA[,c("ENSG00000204642","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000204642,)) + 
  ylab("HLA-F")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-G
```{r}
#ENSG00000204632
scplot<-SubsetA[,c("ENSG00000204632","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000204632,)) + 
  ylab("HLA-G")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DMA
```{r}
#ENSG00000204257
scplot<-SubsetA[,c("ENSG00000204257","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000204257,)) + 
  ylab("HLA-DMA")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DMB
```{r}
#ENSG00000242574
scplot<-SubsetA[,c("ENSG00000242574","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000242574,)) + 
  ylab("HLA-DMB")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DOA
```{r}
#ENSG00000204252
scplot<-SubsetA[,c("ENSG00000204252","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000204252,)) + 
  ylab("HLA-DOA")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```

###HLA-DOB
```{r}
#ENSG00000241106
scplot<-SubsetA[,c("ENSG00000241106","SequencingID","groups")]

#ggplot, with added boxplot and violin shape to visualize density distribution
ggplot(scplot, aes(groups, ENSG00000241106,)) + 
  ylab("HLA-DOB")+
  theme_light()+
  geom_point( aes(colour=groups)) + 
  geom_boxplot(alpha = 0, width = 0.3) + 
  geom_violin(alpha=0.2, width=0.3, aes(fill=groups)
  )


```









